### 2014-01-30 manuscripts/dpp-msbayes/dpp-msbayes-main.tex
Updating DPP manuscript.
Adding Adam's comments to abstract and intro.

\section*{Introduction}
Understanding the processes that generate biodiversity and regulate
community assembly is a major goal of evolutionary biology.
Given the dynamic nature of our planet, a major component of this goal is
understanding how geological and climatic processes can affect the
evolutionary history of whole groups of co-distributed species and their
associated microbiota, including pathogens.
Such large-scale processes predict patterns of shared evolutionary history
across affected populations of organisms, e.g., temporally clustered
divergences across co-distributed species.
At recent timescales, such events should leave a signature in the genetic
variation within and among the affected lineages.
Thus, being able to accurately estimate models of shared evolutionary
history across co-distributed taxa would provide an important
tool for better understanding how regional and global biogeographic
processes affect biodiversity.

@@

The software package \msb implements such an approach in an
approximate-Bayesian model-choice framework \citep{Hickerson2006,Huang2011}.
Recently, however, \citet{Oaks2012} found that \msb can be strongly biased
toward spuriously estimating biogeographically interesting models of highly
clustered divergence times across taxa.
The results of their simulation-based assessment of the method suggest that a
combination of broad, uniformly distributed priors on nuisance parameters and a
``U-shaped'' prior on divergence models are causing the bias.
The model implemented in \msb uses a discrete uniform prior over the number of
possible divergence parameters, which \citet{Oaks2012} show imposes a prior on
divergence models that puts most of the prior mass on models with either very
few or very many divergence times shared across the taxa (see Figure 5 of
\citet{Oaks2012}; for brevity we will refer to this prior as ``U-shaped'').
Furthermore, uniform priors used for most of the model's parameters
result in small marginal likelihoods of parameter-rich models, i.e., models
with many divergence-time parameters \citep{Jeffreys1935, Lindley1957}.
Thus, together, the priors used in \msb will often create a bias toward
models with few divergence-time parameters \citep{Oaks2012}.

>>>>

\section*{Introduction}
Understanding the processes that generate biodiversity and regulate
community assembly is a major goal of evolutionary biology.
Given the dynamic nature of our planet, a major component of this goal is
understanding how geological and climatic processes can affect the
evolutionary history of whole groups of co-distributed species and their
associated microbiota, including pathogens.
Such large-scale processes predict patterns of shared evolutionary history
across affected populations of organisms, e.g., temporally clustered
divergences across co-distributed species.
At recent timescales, such events should leave a signature in the genetic
variation within and among the affected lineages.
Thus, accurately estimating models of shared evolutionary history across
co-distributed taxa provides an important tool for better understanding how
regional and global biogeographic processes affect biodiversity.

@@

The software package \msb implements such an approach in an
approximate-Bayesian model-choice framework \citep{Hickerson2006,Huang2011}.
Given aligned sequence data for \npairs{} pairs of populations, \msb 
estimates the number of divergence events shared among the pairs, the timing of
the events, and the assignment of pairs to the events, while integrating out
uncertainty in demographic parameters and the genealogical history of the
sequences.
% samples over all the possible ways of assigning the divergences between the
% populations of each pair to 1--\npairs{} divergence events, while estimating
% the timing of the events and integrating out uncertainty in demographic
% parameters and the genealogical history of the sequences.
Thus, when the number of pairs is not trivial, the method samples over a large
number of divergence models of differing dimensionality (i.e., all the possible
partitions of \npairs{} pairs to 1--\npairs{} divergence-time parameters), and,
in so doing, estimates the posterior probability of each possible divergence
model.

For priors on divergence-time and demographic parameters, \msb uses continuous
uniform probability distributions
This causes  divergence models with more divergence-time parameters to
integrate over a \emph{much} greater parameter space with low likelihood yet
high prior density, resulting in small marginal likelihoods relative to models
with fewer divergence-time parameters.
Furthermore, \msb uses a discrete uniform prior over the number of divergence
events $1,2,\ldots,\npairs{}$.
Because there are many more possible divergence models for intermediate numbers
of divergence events, this imposes a prior on divergence models that strongly
disfavors such models and puts most of the prior mass on models with either
very few or very many divergence-time parameters (see Figure 5 of
\citet{Oaks2012}; for brevity we will refer to this prior as ``U-shaped'').
Given that models with many divergence-time parameters have small marginal
likelihoods due to the uniform priors on divergence-time and demographic
parameters, the U-shaped prior will strongly favor models with very
few divergence events.
As a result, even when highly informative uniform priors are used for
divergence-time and demographic parameters, \msb will often strongly support
biogeographically interesting models with a small number of divergence events
shared among the taxa when, in reality, divergences were random over broad
timescales (\citet{Oaks2012}).

% Recently, however, \citet{Oaks2012} found that \msb can be strongly biased
% toward spuriously estimating biogeographically interesting models of highly
% clustered divergence times across taxa.
% The results of their simulation-based assessment of the method suggest that a
% combination of broad, uniformly distributed priors on nuisance parameters and a
% ``U-shaped'' prior on divergence models are causing the bias.
% The model implemented in \msb uses a discrete uniform prior over the number of
% possible divergence parameters, which \citet{Oaks2012} show imposes a prior on
% divergence models that puts most of the prior mass on models with either very
% few or very many divergence times shared across the taxa (see Figure 5 of
% \citet{Oaks2012}; for brevity we will refer to this prior as ``U-shaped'').
% Furthermore, uniform priors used for most of the model's parameters
% result in small marginal likelihoods of parameter-rich models, i.e., models
% with many divergence-time parameters \citep{Jeffreys1935, Lindley1957}.
% Thus, together, the priors used in \msb will often create a bias toward
% models with few divergence-time parameters \citep{Oaks2012}.

### 2014-01-30 manuscripts/dpp-msbayes/dpp-msbayes.tex
Updating DPP manuscript.
Adding Adam's comments to abstract and intro.

{\sffamily
    \noindent\textbf{ABSTRACT} \\
    \noindent To understand the processes that generate biodiversity, it is
    important to account for large-scale processes that affect the evolutionary
    history of groups of co-distributed populations of organisms.
    Such events predict temporally clustered divergences times, a pattern
    that can be estimated using genetic data from co-distributed species.
    We introduce a new approximate-Bayesian model for comparative
    phylogeographical model-choice that estimates the temporal distribution of
    divergences across taxa from multi-locus DNA sequence data.
    Our model is an extension of that implemented in \msb.
    By reparameterizing the model, introducing more flexible priors on
    demographic and divergence-time parameters, and implementing a
    non-parametric Dirichlet-process prior over divergence models, we improve
    the robustness, accuracy, and power of the method for estimating shared
    evolutionary history across taxa.
    Our results demonstrate that the tendency of the \msb model to spuriously
    support models of clustered divergences is caused by a combination of (1)
    uniform priors on nuisance parameters hindering the marginal likelihoods of
    models with more divergence-time parameters, and (2) a prior on divergence
    models that disfavors models with intermediate numbers of divergence-time
    parameters.
    The new method addresses both of these issues and yields more robust
    estimates of posterior uncertainty, and thus greatly reduces the tendency
    of the model to incorrectly estimate biogeographically interesting models
    with strong support.

>>>>

{\sffamily
    \noindent\textbf{ABSTRACT} \\
    \noindent To understand the processes that generate biodiversity, it is
    important to account for large-scale processes that affect the evolutionary
    history of groups of co-distributed populations of organisms.
    Such events predict temporally clustered divergences times, a pattern
    that can be estimated using genetic data from co-distributed species.
    We introduce a new approximate-Bayesian model for comparative
    phylogeographical model-choice that estimates the temporal distribution of
    divergences across taxa from multi-locus DNA sequence data.
    Our model is an extension of that implemented in \msb.
    By reparameterizing the model, introducing more flexible priors on
    demographic and divergence-time parameters, and implementing a
    non-parametric Dirichlet-process prior over divergence models, we improve
    the robustness, accuracy, and power of the method for estimating shared
    evolutionary history across taxa.
    Our results demonstrate the improved performance of the new method is due
    to (1) more appropriate priors on divergence-time and demographic
    parameters that avoid prohibitively small marginal likelihoods for models
    with more divergence events,
    and (2) the Dirichlet-process providing a flexible prior on divergence
    histories that does not strongly disfavor models with intermediate numbers
    of divergence events.
    % Our results demonstrate that the tendency of the \msb model to spuriously
    % support models of clustered divergences is caused by a combination of (1)
    % uniform priors on nuisance parameters hindering the marginal likelihoods of
    % models with more divergence-time parameters, and (2) a prior on divergence
    % models that disfavors models with intermediate numbers of divergence-time
    % parameters.
    The new method yields more robust estimates of posterior uncertainty, and
    thus greatly reduces the tendency of the model to incorrectly estimate
    biogeographically interesting models with strong support.
